---
title: "Maharjan_He_Final_Project"
author: "Meelisha Maharjan and Xinchen He"
date: "2022-11-24"
output: powerpoint_presentation
---

## Import libaries

```{r}
# install.packages("dataRetrieval")
library(dataRetrieval)
suppressMessages(library(tidyverse))
library(dataRetrieval)
library(dplyr)
library(ggplot2)
```

## Data Collection

- Define a function to scrap the time series of the streamflow (discharge).
```{r}
# Define function to get discharge timeseries
get_discharge_data <- function(site_no, q_code, start, end) {
  
  # Build the url link
  url <- paste0('https://waterdata.usgs.gov/nwis/dv?cb_',
        q_code,
        '=on&format=html&site_no=',
        site_no,
        '&legacy=&re[…]module=sw&period=&begin_date=',
          start,
        '&end_date=', 
        end)
  q_data <- readLines(url)
  
  # Select lines which contains the data
  table_pattern <- '<tr align="center"><td nowrap="nowrap">'
  q_table <- str_subset(q_data, table_pattern)
  
  # Extract dates and flow(ft3/s) from each line
  dates = as.Date(unlist(str_match_all(q_table, "[0-9]*/[0-9]*/[0-9]*")), "%m/%d/%Y")
  flows = unlist(str_match_all(q_table, "<span>.*</span>"))
  flows = as.numeric(str_replace_all(flows, "</?span>", ""))
  
  # Put dates and flow together as a matrix
  Q_df = data.frame(
    Date = dates,
    Qobs = flows
  )
  return(Q_df)
}
```

- Define a function to scrap the drainge area of the gauge station
```{r}
get_drainage_area = function(site_no){
   # Scrape the drainage area
   url2 <- paste0('https://waterdata.usgs.gov/nwis/inventory/?site_no=',
        site_no,
        '&agency_cd=USGS')
   d_data <- readLines(url2)
   d_pattern <- "^.*Drainage area"
   d_line <- str_subset(d_data,d_pattern)
   darea <- unique(unlist(str_extract_all(d_line, "[0-9]*\\.?[0-9]*")))[2]
   darea <- as.numeric(darea)
   return(darea)
}

```

- Validate the scraped data with the data provided by the USGS `dataRetrieval`

```{r}
# Specify the inputs: gauge number, q_code is fixed, start date and end date
site_no <- '08324000'
q_code <- '00060'
start <- '2000-01-01'
end <- '2020-12-31'
# Use the function we have to scrape
q_df <- get_discharge_data(site_no, q_code, start, end)

# Get authentic streamflow data using the package
validate_q <- readNWISdv(siteNumber = site_no, parameterCd = q_code,
  start, end)

# Calculate the error between scraped data and the true data
scrape_error = q_df$Qobs - validate_q$X_00060_00003

plot(q_df$Date, scrape_error)
```

- Get streamflow time series and drainge area of the gauges in Massachusetts (TAKE A LOT OF TIME, DO NOT RUN IF YOU HAVE DATA EXISTED IN THE FOLDER!!!)

```{r}
# Load the site list
# site_list <- readLines("gauges_Mass.txt")
# create an empty list to save drainage area
# da_area <- numeric(length(site_list))#

# counter = 1
# for (site_no in site_list){
#   try({
#     da_area[counter] = get_drainage_area(site_no)
#     })
#   counter <- counter + 1
#   print(site_no)
# }

# Note, not all the sites have valid data in the date range
# counter = 1
# for (site_no in site_list){
#   print(paste('Gauge ',site_no,' Scraping start'))
#   
#   # scrape discharge data
#   q_df <- try({
#     get_discharge_data(site_no, q_code, start, end)
#   }, silent = TRUE)
#   
#   # scrape the draiange area
#   try({
#     da_area[counter] <- get_drainage_area(site_no)
#     })
#   
#   # Save the streamflow
#   try({
#     write.csv(q_df, paste0("streamflows/", site_no,".csv"), row.names = FALSE)
#   }, silent = TRUE)
#   
#   counter = counter + 1
#   print(paste('Gauge ',site_no,' Done'))
# }
# 
# # Save the drainge area
# write.csv(data.frame(
#   site_no <- site_list,
#   drainage_area <- da_area
# ), "drainage_area.txt", row.names = FALSE)
```

## Calculate Streamflow Statistics and Visualizations

```{r}
# Note: not all the site have a validate
# 0 drainage area meant no data avaliable

gauges_area = read.csv("drainage_area.txt", colClasses = c("character", "numeric"))

# Filter gauges which do not have the drainage area data
gauges_area <- gauges_area %>%
                filter(drainage_area > 0)
                
gauges_area
```
- Import streamflows and calculate streamflow statistics (drop site with no drainage area and less than 2 years observations)

```{r}
# Note: not all the gauges have enough streamflow data
site_list <- character()
Q5s <- numeric()
Q50s <- numeric()
Q95s <- numeric()

for (site_no in gauges_area$site_no){
  flow_df <- read.csv(paste0("streamflows/",
                            site_no, ".csv")
                    )
  # check how many data in the data
  if (nrow(flow_df) <= 15 * 365){
    # if less 2 years data were scraped
    # go to next gauge
    print(paste0("Less than 2 years in the site:", site_no))
  }else{
    # more than 2 years data were scraped
    # the gauge would be saved into the site_list
    
    # save the site no
    site_list = c(site_list, site_no)
    
    # Calculate the statistics
    Q5 = unname(quantile(x = flow_df$Qobs, probs = c(.95), na.rm = TRUE))
    Q50 = unname(quantile(x = flow_df$Qobs, probs = c(.5), na.rm = TRUE))
    Q95 = unname(quantile(x = flow_df$Qobs, probs = c(.05), na.rm = TRUE))
    
    # save the statistics
    Q5s = c(Q5s, Q5)
    Q50s = c(Q50s, Q50)
    Q95s = c(Q95s, Q95)
  }
}
```
- Build a new dataframe which contains the site_list, Q5, Q95, Q50, drainage_area
```{r}
# The dataframe is based on the gauge area dataframe
Q_A_df = gauges_area %>%
            filter(site_no %in% site_list) %>%
            mutate(Q5 = Q5s) %>%
            mutate(Q50 = Q50s) %>%
            mutate(Q95 = Q95s)
Q_A_df = drop_na(Q_A_df)
Q_A_df
```

## Build a regression between drainage_are and Q5, Q50, Q95

- First, preview the relationship between the drainage area (X) and Q5(Y1), Q50(Y2) and Q95(Y3)

```{r}
# they are linear related when they were logged
ggplot(Q_A_df, aes(log(drainage_area), y = log(Q5))) + geom_point()
```

```{r}
ggplot(Q_A_df, aes(log(drainage_area), y = log(Q95))) + geom_point()
```
- Define a $\psi$ function to optimize the linear function.
```{r}
psi <- function(x, c = 1) {
  return (
    ifelse(
      abs(x) > c,
      2 * c * abs(x) - c^2,
      x^2
    )
  )
}

mean_psi_regression_q5 <- function(b, data, c = 1) {
  b1 = b[1]
  b2 = b[2]
  Y = data$Q5
  X = data$drainage_area
  return( mean ( psi( log(Y) - (b1 + b2 * log(X) ), c ) ) )
}

mean_psi_regression_q95 <- function(b, data, c = 1) {
  b1 = b[1]
  b2 = b[2]
  Y = data$Q95
  X = data$drainage_area
  return( mean ( psi( log(Y) - (b1 + b2 * log(X) ), c ) ) )
}

```

- Use `optim` function to get the best values of the coefficients
```{r}
best_q5 = optim(par = c(0,0), mean_psi_regression_q5, data = Q_A_df, c= 1)
(best_q5)

best_q95 <- optim(par = c(0,0), mean_psi_regression_q95, data = Q_A_df, c = 1)
(best_q95)
```
- Bootstrap to get the distribution of the coefficients.
```{r}
# Bootstrap
n <- nrow(Q_A_df)
M <- 10000
b0_bootstrap_sample_q5 <- numeric(M)
b1_bootstrap_sample_q5 <- numeric(M)

b0_bootstrap_sample_q95 <- numeric(M)
b1_bootstrap_sample_q95 <- numeric(M)

for (m in 1:M) {
  bootstrap_sample_idx <- sample(n, replace = TRUE)
  bstp <- Q_A_df[bootstrap_sample_idx, ]
  b_psi_GD_q5 <- optim(c(0, 0), mean_psi_regression_q5, data = bstp)
  b_psi_GD_q95 <- optim(c(0, 0), mean_psi_regression_q95, data = bstp)
  b0_bootstrap_sample_q5[m] <- b_psi_GD_q5$par[1]
  b1_bootstrap_sample_q5[m] <- b_psi_GD_q5$par[2]
  b0_bootstrap_sample_q95[m] <- b_psi_GD_q95$par[1]
  b1_bootstrap_sample_q95[m] <- b_psi_GD_q95$par[2]
  
}
```

- q5 bootstrap samples plot
```{r}
q5_plt1 <- ggplot(data.frame(b0_bootstrap_sample_q5), aes(b0_bootstrap_sample_q5)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.2)
q5_plt2 <- ggplot(data.frame(b1_bootstrap_sample_q5), aes(b1_bootstrap_sample_q5)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.2)
mean(b0_bootstrap_sample_q5)
mean(b1_bootstrap_sample_q5)
q5_plt1
q5_plt2
```
- q95 bootstrap samples plot
```{r}
q95_plt1 <- ggplot(data.frame(b0_bootstrap_sample_q95), aes(b0_bootstrap_sample_q95)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.2)
q95_plt2 <- ggplot(data.frame(b1_bootstrap_sample_q95), aes(b1_bootstrap_sample_q95)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.2)
mean(b0_bootstrap_sample_q95)
mean(b1_bootstrap_sample_q95)
q95_plt1
q95_plt2
```


- Obtain q5 confidence interval and plot
```{r}
alpha <- 0.05
q5_bootstrap_CI_0 <- quantile(b0_bootstrap_sample_q5, c(alpha/2, 1 - alpha/2))
q5_bootstrap_CI_1 <- quantile(b1_bootstrap_sample_q5, c(alpha/2, 1 - alpha/2))
q5_plt1 + geom_vline(xintercept = q5_bootstrap_CI_0, col = "blue")
q5_plt2 + geom_vline(xintercept = q5_bootstrap_CI_1, col = "blue")
```


- Obtain q95 confidence interval and plot
```{r}
alpha <- 0.05
q95_bootstrap_CI_0 <- quantile(b0_bootstrap_sample_q95, c(alpha/2, 1 - alpha/2))
q95_bootstrap_CI_1 <- quantile(b1_bootstrap_sample_q95, c(alpha/2, 1 - alpha/2))
q95_plt1 + geom_vline(xintercept = q95_bootstrap_CI_0, col = "blue")
q95_plt2 + geom_vline(xintercept = q95_bootstrap_CI_1, col = "blue")
```

```{r}
new_df = Q_A_df %>%
  mutate(log_darea = log(drainage_area)) %>%
  mutate(logQ5 = log(Q5)) %>%
  mutate(logQ95 = log(Q95)) %>%
  # mutate(logQ5_predicted_mean_optim =  best$par[1] + best$par[2] * log_darea) %>%
  mutate(logQ5_predicted_mean =  mean(b0_bootstrap_sample_q5) + mean(b1_bootstrap_sample_q5) * log_darea)  %>%
  # mutate(logQ95_predicted_mean_optim =  best$par[1] + best$par[2] * log_darea) %>%
  mutate(logQ95_predicted_mean =  mean(b0_bootstrap_sample_q95) + mean(b1_bootstrap_sample_q95) * log_darea) %>%
  mutate(logQ5_predicted_25 =  mean(q5_bootstrap_CI_0[1]) + mean(q5_bootstrap_CI_1[1]) * log_darea) %>%
  mutate(logQ5_predicted_975 =  mean(q5_bootstrap_CI_0[2]) + mean(q5_bootstrap_CI_1[2]) * log_darea) %>%
  mutate(logQ95_predicted_25 =  mean(q95_bootstrap_CI_0[1]) + mean(q95_bootstrap_CI_1[1]) * log_darea) %>%
  mutate(logQ95_predicted_975 =  mean(q95_bootstrap_CI_0[2]) + mean(q95_bootstrap_CI_1[2]) * log_darea)

q5_reg_plots <- ggplot(data = new_df, aes(x = log_darea, y = logQ5)) +
            geom_point() + 
            # geom_line(data = new_df, aes(x = log_darea, y = logQ95_predicted_mean_optim), color = 'red') + 
            geom_line(data = new_df, aes(x = log_darea, y = logQ95_predicted_mean), color = 'green') +
            geom_line(data = new_df, aes(x = log_darea, y = logQ95_predicted_25), color = 'blue') +
            geom_line(data = new_df, aes(x = log_darea, y = logQ95_predicted_975), color = 'black')

q95_reg_plots <- ggplot(data = new_df, aes(x = log_darea, y = logQ95)) +
            geom_point() + 
            # geom_line(data = new_df, aes(x = log_darea, y = logQ95_predicted_mean_optim), color = 'red') + 
            geom_line(data = new_df, aes(x = log_darea, y = logQ5_predicted_mean), color = 'green') +
            geom_line(data = new_df, aes(x = log_darea, y = logQ5_predicted_25), color = 'blue') +
            geom_line(data = new_df, aes(x = log_darea, y = logQ5_predicted_975), color = 'black')

```


```{r}
q5_reg_plots
```


